### 2024-03-09 outputs/paper.qmd
graphs and cleaning

---
title: "Joe Biden won popular votes in 2024 US Presidential Election"
format: pdf
ate: "9 March 2024"
date-format: "D MMMM YYYY"
number-sections: true
thanks: "Code and data in this analysis is available at: "
toc: true

---
```{r}
#| include: false
# install.packages('dplyr')
# install.packages('ggplot2')
# install.packages('knitr')
# install.packages('statebins')
# install.packages('modelsummary')
# install.packages('broom')
# install.packages("here")
# install.packages("kableExtra")

library(dplyr)
library(ggplot2)
library(knitr)
library(statebins)
library(modelsummary)
library(broom)
library(readr)
library(here)
library(kableExtra)
```

```{r}
#| include: false
data_ipums_clean <-
  read_csv(
    file = here("inputs/IPUMS/cleaned_population_data.csv"),
    show_col_types = FALSE
  )

data_app_clean <-
  read_csv(
    file = here("inputs/APP/cleaned_sample_data.csv"),
    show_col_types = FALSE
  )

data_app_clean_20 <-
  read_csv(
    file = here("inputs/APP/cleaned_sample_data_20.csv"),
    show_col_types = FALSE
  )

statevote2020 <-
  read_csv(
    file = here("inputs/Kaggle/cleaned_actual_data_20.csv"),
    show_col_types = FALSE
  )

data_app_raw <-
  read_csv(
    file = here("inputs/APP/raw_sample_data.csv"),
    show_col_types = FALSE
  )
```

```{r}
#| include: false
race <- data_app_clean |> 
  group_by(race) |> 
  summarise(n = n()) |> 
  mutate(pct = n / sum(n),
         type = "pre-strat",
         group = "race") |> 
  rename(level = race)

gender <- data_app_clean |> 
  group_by(gender) |> 
  summarise(n = n()) |> 
  mutate(pct = n / sum(n),
         type = "pre-strat",
         group = "gender") |> 
  rename(level = gender)

education <- data_app_clean |> 
  group_by(education) |> 
  summarise(n = n()) |> 
  mutate(pct = n / sum(n),
         type = "pre-strat",
         group = "education") |> 
  rename(level = education)

marital <- data_app_clean |> 
  group_by(marital) |> 
  summarise(n = n()) |> 
  mutate(pct = n / sum(n),
         type = "pre-strat",
         group = "marital") |> 
  rename(level = marital)

age <- data_app_clean |> 
  group_by(age) |> 
  summarise(n = n()) |> 
  mutate(pct = n / sum(n),
         type = "pre-strat",
         group = "age") |> 
  rename(level = age)

statename <- data_app_clean |> 
  group_by(statename) |> 
  summarise(n = n()) |> 
  mutate(pct = n / sum(n),
         type = "pre-strat",
         group = "statename") |> 
  rename(level = statename)
```

```{r}
#| include: false
race_post <- data_ipums_clean |> 
  group_by(race) |> 
  summarise(n = n()) |> 
  mutate(pct = n / sum(n),
         type = "post-strat",
         group = "race") |> 
  rename(level = race)

gender_post <- data_ipums_clean |> 
  group_by(gender) |> 
  summarise(n = n()) |> 
  mutate(pct = n / sum(n),
         type = "post-strat",
         group = "gender") |> 
  rename(level = gender)

education_post <- data_ipums_clean |> 
  group_by(education) |> 
  summarise(n = n()) |> 
  mutate(pct = n / sum(n),
         type = "post-strat",
         group = "education") |> 
  rename(level = education)

marital_post <- data_ipums_clean |> 
  group_by(marital) |> 
  summarise(n = n()) |> 
  mutate(pct = n / sum(n),
         type = "post-strat",
         group = "marital") |> 
  rename(level = marital)

age_post <- data_ipums_clean |> 
  group_by(age) |> 
  summarise(n = n()) |> 
  mutate(pct = n / sum(n),
         type = "post-strat",
         group = "age") |> 
  rename(level = age)

statename_post <- data_ipums_clean |> 
  group_by(statename) |> 
  summarise(n = n()) |> 
  mutate(pct = n / sum(n),
         type = "post-strat",
         group = "statename") |> 
  rename(level = statename)
```

```{r}
#| include: false
variable1 <- rbind(gender, gender_post, marital, marital_post, race, race_post, age, age_post)
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-demographics1
#| fig-cap: Voter Demographics (Age, Gender, Marital Status, Race)
variable1 |>
  ggplot(aes(as.factor(level), pct, group = as.factor(type), linetype = as.factor(type))) + 
  geom_line() + 
  facet_wrap(~ group, nrow = 2, scales = "free") + 
  theme(axis.text.x = element_text(angle = 45, size = 7, hjust = 1, vjust = 1)) +
  labs(x = "Category", y = "Proportion", linetype = "Data Set") + 
  scale_y_continuous(labels = scales::percent) +
  geom_hline(yintercept = 0.5, alpha = 0.3)
```

```{r}
#| include: false
variable2 <- rbind(education, education_post, statename, statename_post)
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-demographics2
#| fig-cap: Voter Demographics (Education, State of Residence)
variable2 |> 
  ggplot(aes(level, pct, group = as.factor(type), linetype = as.factor(type))) + 
  geom_line() + 
  facet_wrap(~ group, nrow = 2, scales = "free") + 
  theme(axis.text.x = element_text(angle = 45, size = 5, hjust = 1)) +
  labs(x = "Category", y = "Proportion", linetype = "Data Set",
       title = "Voter Demographics (Education, State of Residence)") +
  scale_y_continuous(labels = scales::percent)
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-survey_vote
#| fig-cap: Proportion of the Two Parties in the Survey
data_app_clean |>
  group_by(pid3) |>
  summarise(n = n()) |>
  mutate(Proportion = round(n/sum(n)*100)) |>
  ggplot(aes(x = pid3, y = n)) +
  geom_bar(stat = "identity", fill = "grey", color = "black") +
  geom_text(aes(label = paste0(Proportion, "%")), vjust = -0.5, color = "black") +
  labs(x = "Party",
       y = "Number of Respondents") + 
  scale_y_continuous(limits = c(0, 10000)) +
  theme_minimal()
```

```{r}
#| include: false
survey_map <- data_app_clean |>
  group_by(statename, vote_biden) |>
  summarise(n = n()) |>
  group_by(statename) |>
  mutate(prop = n/sum(n)) |>
  filter(vote_biden == "1")
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-sample_map
#| fig-cap: Support for Joe Biden and Donald Trump by State in the Sample
survey_map |>
  ggplot(aes(fill = prop, state = statename)) + 
  geom_statebins() + 
  scale_fill_gradient2(low = "#971116", high = "#005D94", mid = "white", midpoint = 0.5) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion of \nSupporting Biden")
```

```{r}
#| include: false
model <- glm(vote_biden ~ race + gender + education + statename + marital + age, 
             data = data_app_clean, family = binomial())
```

```{r}
#| include: false
saveRDS(
  model,
  file = "us_election_model_mrp.rds"
)
```

```{r}
#| include: false
modelsummary(model)
coefficients <- broom::tidy(model, conf.int = TRUE)
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-coefficients
#| tbl-cap: Coefficients of the Voting Model
coefficients |> kable(digits = 2)
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-coefficients_model
#| fig-cap: Estimates of Coefficients in the Model
coefficients |> 
  ggplot(aes(x = estimate, y = term)) + 
  geom_point() +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), color = "skyblue3", position = position_dodge(width = 0.2)) +
  labs(x = "Estimates", y = "Coefficients") +
  theme_bw() + 
  theme(axis.text.y = element_text(size = 4.5))
```

```{r}
#| include: false
proportion_data <- data_ipums_clean |>
  group_by(statename, race, age, gender, marital, education) |>
  summarise(n = n()) |> 
  group_by(statename) |>
  mutate(prop = n/sum(n))

proportion_data$estimate <- predict.glm(model, newdata = proportion_data, type = "response")
proportion_data <- proportion_data |> mutate(num_voters = n*estimate)

errors <- predict.glm(model, newdata = proportion_data, type = "response", se.fit = T)
upper_e <- errors$fit + errors$se.fit
lower_e <- errors$fit - errors$se.fit
proportion_error <- cbind(proportion_data, lower_e, upper_e) |> 
  rename("lower_bound" = ...11, "upper_bound" = ...12)
```

```{r}
#| include: false
proportion_error <- proportion_error |>
  mutate(lower_estimate = n*lower_bound, upper_estimate = n*upper_bound)

vote_for_biden <- proportion_error |> mutate(biden_predict_prop = estimate* prop,
                                       biden_predict_prop_lower = lower_bound * prop,
                                       biden_predict_prop_upper = upper_bound * prop) |>
  group_by(statename) |>
  summarise(biden_predict = sum(biden_predict_prop),
            biden_lower = sum(biden_predict_prop_lower),
            biden_upper = sum(biden_predict_prop_upper))
```

```{r}
#| include: false
data1 <- data_app_clean |>
  group_by(statename, vote_biden) |>
  summarise(n = n()) |>
  mutate(prop = n / sum(n))

data2 <- inner_join(data1, vote_for_biden, by = "statename") |> filter(vote_biden == '1')

```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-sample_and_population
#| fig-cap: Proportion of Voters in Each State Voting for Biden
data2 |>
  ggplot(aes(x = statename, y = biden_predict * 100)) +
  geom_point(aes(color = "Biden Prediction"), size = 2) + 
  geom_line(aes(group = 1), linetype = "solid", size = 0.8) +
  geom_ribbon(aes(group = 1, ymin = biden_lower * 100, ymax = biden_upper * 100), alpha = 0.2, fill = "blue3") +
  geom_line(aes(x = statename, y = prop * 100, group = 2), 
            linetype = "dashed", color = "red2", size = 0.8) +
  geom_point(data = data2, aes(x = statename, y = prop * 100, color = "Population"), size = 2) +
  geom_hline(aes(yintercept = 40), alpha = 0.5, linetype = "dotted", color = "gray") +
  geom_hline(aes(yintercept = 60), alpha = 0.5, linetype = "dotted", color = "gray") +
  geom_hline(aes(yintercept = 80), alpha = 0.5, linetype = "dotted", color = "gray") +
  geom_hline(aes(yintercept = 50), alpha = 0.6, color = "blue4") +
  scale_color_manual(name = "Data", values = c("Biden Prediction" = "black", "Population" = "red2"), labels = c("Population Prediction", "Survey Sample")) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 55, size = 6, hjust = 1, vjust = 1),
        legend.position = "bottom",
        legend.title = element_text(size = 10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  labs(x = "State", y = "Voting Percentage for Biden")
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-population_map
#| fig-cap: Predicted support for Joe Biden and Donald Trump by State
vote_for_biden |>
  ggplot(aes(fill = biden_predict, state = statename)) + 
  geom_statebins() + 
  scale_fill_gradient2(low = "#971116", high = "#005D94",
                       mid = "white", midpoint = 0.5) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion of \nSupporting Biden")
```

```{r}
#| include: false
final_data <- data2 |>
  inner_join(data_ipums_clean, by = c("statename" = "statename"))

total_proportion <- sum(final_data$biden_predict * (final_data$n / sum(final_data$n)))
total_proportion
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-survey_votes_2020
#| tbl-cap: Votes in 2020 from the Survey Respondents
data_app_raw |>
  group_by(presvote20post) |>
  summarize(count = n()) |>
  mutate(proportion = count / sum(count) * 100) |> 
  rename('Proportion (%)' = proportion, 
         'Candidates' = presvote20post, 
         'Number of Votes' = count) |> 
  kable()
```

```{r}
#| include: false
pid3_levels <- c("Republican", "Democrat")
```

```{r}
#| label: fig-balanced_two_compare
#| fig-cap: Distribution of Supporting Parties Now and Votes in 2020 from the Survey Respondents
#| echo: false
#| layout-ncol: 2
#| warning: false
#| message: false
#| fig-subcap: ["Parties Respondents Support Now", "Respondents' votes in 2020"]
data_app_clean_20 |> mutate(pid3 = factor(pid3, levels = pid3_levels)) |>
  ggplot(aes(x = pid3)) + 
  geom_bar(fill = "grey", color = "black") +
  geom_text(stat = "count", aes(label = ..count..), vjust = 2) +
  labs(x = "Party", y = "Number of Counts") +
  theme_minimal()

data_app_clean_20 |>
  ggplot(aes(x = vote2020)) +
  geom_bar(fill = "grey", color = "black") +
  geom_text(stat = "count", aes(label = ..count..), vjust = 2) +
  labs(x = "Candidate in 2020", y = "Number of Counts") +
  theme_minimal()
```

```{r}
#| include: false
survey_map_20 <- data_app_clean_20 |>
  group_by(statename, vote_biden) |>
  summarise(n = n()) |>
  group_by(statename) |>
  mutate(prop = n / sum(n)) |>
  filter(vote_biden == "1")
```

```{r}
#| label: fig-actual_survey_compare
#| fig-cap: Distribution of prices of Paris Airbnb rentals in December 2023
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 2
#| fig-subcap: ["2020 Actual Voting Results by States", "2020 Voting Results From Balanced Survey Respondentsby by States"]

# Actual votes in 2020
statevote2020 |> 
  ggplot(aes(fill = support_biden_prop, state = state)) + 
  geom_statebins() + 
  scale_fill_gradient2(low = "#971116", high = "#005D94",
                       mid = "white", midpoint = 0.5) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion of \nSupporting Biden")

# Survey votes in 2020
survey_map_20 |>
  ggplot(aes(fill = prop, state = statename)) + 
  geom_statebins() + 
  scale_fill_gradient2(low = "#971116", high = "#005D94", mid = "white", midpoint = 0.5) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion of \nSupporting Biden")
```

```{r}
#| include: false
model2 <- glm(vote_biden ~ race + gender + education + statename + marital + age, 
             data = data_app_clean_20, family = binomial())

coefficients_20 <- broom::tidy(model2, conf.int = TRUE)
```

```{r}
#| include: false
saveRDS(
  model2,
  file = "us_election_model_balanced_mrp.rds"
)
```

```{r}
#| include: false
proportion_data_20 <- data_ipums_clean |>
  group_by(statename, race, age, gender, marital, education) |>
  summarise(n = n()) |> 
  group_by(statename) |>
  mutate(prop = n/sum(n))

proportion_data_20$estimate <- predict.glm(model2, newdata = proportion_data_20, type = "response")
proportion_data_20 <- proportion_data_20 |> mutate(num_voters = n*estimate)

errors_20 <- predict.glm(model2, newdata = proportion_data_20, type = "response", se.fit = T)
upper_e_20 <- errors_20$fit + errors_20$se.fit
lower_e_20 <- errors_20$fit - errors_20$se.fit
proportion_error_20 <- cbind(proportion_data_20, lower_e_20, upper_e_20) |> 
  rename("lower_bound" = ...11, "upper_bound" = ...12)

proportion_error_20 <- proportion_error_20 |>
  mutate(lower_estimate_20 = n*lower_bound, upper_estimate_20 = n*upper_bound)
```

```{r}
#| include: false
vote_for_biden_20 <- proportion_error_20 |> mutate(biden_predict_prop_20 = estimate* prop,
                                       biden_predict_prop_lower_20 = lower_bound * prop,
                                       biden_predict_prop_upper_20 = upper_bound * prop) |>
  group_by(statename) |>
  summarise(biden_predict_20 = sum(biden_predict_prop_20),
            biden_lower_20 = sum(biden_predict_prop_lower_20),
            biden_upper_20 = sum(biden_predict_prop_upper_20))
```

```{r}
#| include: false
data3 <- data_app_clean_20 |>
  group_by(statename, vote_biden) |>
  summarise(n = n()) |>
  mutate(prop = n / sum(n))

data4 <- inner_join(data3, vote_for_biden_20, by = "statename") |> filter(vote_biden == '1')
```

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: fig-balanced_map
#| fig-cap: Predicted support for Joe Biden and Donald Trump by State (Using Balanced Survey Results)
vote_for_biden_20 |>
  ggplot(aes(fill = biden_predict_20, state = statename)) + 
  geom_statebins() + 
  scale_fill_gradient2(low = "#971116", high = "#005D94",
                       mid = "white", midpoint = 0.5) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion of \nSupporting Biden")
```

```{r}
#| include: false
final_data_20 <- data4 |>
  inner_join(data_ipums_clean, by = c("statename" = "statename"))

total_proportion_20 <- sum(final_data_20$biden_predict_20 * (final_data_20$n / sum(final_data_20$n)))
total_proportion_20
```